---
title: "College Scorecard"
author: "Anh Do"
date: "4/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
```

```{r loading-package}
library(tidyverse)
library(maps)
library(highcharter)
library(plotly)
library(ggrepel)
library(modelr)
library(ggalt)
library("RColorBrewer")
library(wesanderson)
library(ggthemes)
library(gganimate)
```

```{r global-variables}
excluded_states <- c("AK", "HI", "GU", "PR", "AS", "DC", "FM")
```


```{r extra-map-data}
us_map_data <- read_csv("https://simplemaps.com/static/data/us-cities/uscitiesv1.5.csv")

us_map_data <- us_map_data %>% rename(
    CITY = city,
    STATENAME = state_name
)
```


```{r}
# data_files <- list.files(path = "~/Downloads/CollegeScorecard_Raw_Data", pattern =  "*.csv", full.names = TRUE)
# 
# 
# data_by_years = tibble(
#     UNITID = character(),
#     CITY = character(),
#     STABBR = character(),
#     UNEMP_RATE = double(),
#     YEAR = integer()
# )
# 
# # temp <- read_csv(data_files[14:14])
# 
# years = seq(1997, 2017)
# 
# for (i in 2:length(years)) {
#     file = data_files[i]
#     print(file)
#     year = years[i]
#     data_by_years <- rbind(data_by_years, read_csv(file) %>% 
#               select(UNITID, CITY, STABBR, UNEMP_RATE) %>% 
#               mutate(UNEMP_RATE = as.numeric(UNEMP_RATE),
#                      YEAR = year))
# }
# 
# data_by_years <- data_by_years %>%
#     rename(state_id = STABBR)
# 
# processed_data_by_years <- left_join(data_by_years, us_map_data %>% 
#                                     select(CITY, state_id, county_name, lat, lng), 
#                                     by = c("CITY", "state_id")) 
# 
# processed_data_by_years <- add_column(processed_data_by_years, 
#            STATENAME = state.name[match(processed_data_by_years$state_id, state.abb)], 
#            .after = "state_id")
# 
# processed_data_by_years <- processed_data_by_years %>%
#     filter(!state_id %in% excluded_states)
```


```{r}
theme_set(theme_light())
# theme_set(my_theme)
```

```{r}
college_data_2017 <- read_csv("~/Downloads/CollegeScorecard_Raw_Data/MERGED2016_17_PP.csv") %>%
    select(UNITID, CITY, STABBR, GRAD_DEBT_MDN_SUPP)

college_data_2007 <- read_csv("~/Downloads/CollegeScorecard_Raw_Data/MERGED2006_07_PP.csv") %>%
    select(UNITID, CITY, STABBR, GRAD_DEBT_MDN_SUPP)

college_data_2007_2017 <- inner_join(college_data_2007, college_data_2017, by = c("UNITID", "CITY", "STABBR"))

college_data_2007_2017 <- add_column(college_data_2007_2017, 
                           STATENAME = state.name[match(college_data_2007_2017$STABBR, state.abb)], 
                           .after = "STABBR")

college_data_2007_2017 <- college_data_2007_2017 %>%
    mutate_at(vars(starts_with("GRAD_DEBT")), as.numeric)

college_data_2007_2017_summarise <- college_data_2007_2017 %>%
    group_by(STATENAME) %>%
    summarise(debt_2007 = median(GRAD_DEBT_MDN_SUPP.x, na.rm = TRUE),
              debt_2017 = median(GRAD_DEBT_MDN_SUPP.y, na.rm = TRUE)) %>%
    filter(!is.na(STATENAME)) 

college_data_2007_2017_summarise %>%
    mutate(STATENAME = fct_reorder(STATENAME, abs(debt_2017))) %>%
    ggplot(aes(x=debt_2007, xend=debt_2017, y=STATENAME)) + 
    geom_dumbbell(color="#a3c4dc", 
                  size=2, 
                  colour_x = "red",
                  colour_xend = "blue") + 
    # geom_label(aes(label = abs(debt_2017 - debt_2007)), hjust = -5) +
    scale_x_continuous(labels = scales::dollar_format(), breaks = seq(0, 30000, 5000)) + 
    annotate("text", label = "2007", x = 16000, y = 49.8, color = "red", size = 8) +
    annotate("text", label = "2017", x = 26000, y = 49.8, color = "blue", size = 8) +
    expand_limits(x = c(5000, 30000)) +
    labs(x=NULL, 
         y=NULL,
         title="How Are Debts in Each State?",
         subtitle = "Year 2007 and Year 2017 (Ordered by highest Debt in 2017)") +
   theme(title = element_text(size = 20),
        axis.text.x = element_text(face = "bold", color = "black", size = 15),
        axis.text.y = element_text(color = "black", size = 15),
        legend.text = element_text(color = "black", size = 15),
        panel.grid.minor=element_blank(),
        panel.grid.major.y=element_blank(),
        panel.grid.major.x=element_line(),
        axis.ticks=element_blank(),
        legend.position="top",
        panel.border=element_blank())

# reorder by difference
college_data_2007_2017_summarise %>%
    mutate(STATENAME = fct_reorder(STATENAME, abs(debt_2017 - debt_2007))) %>%
    ggplot(aes(x=debt_2007, xend=debt_2017, y=STATENAME)) + 
    geom_dumbbell(color="#a3c4dc", 
                  size=2, 
                  colour_x = "red",
                  colour_xend = "blue") + 
    # geom_label(aes(label = abs(debt_2017 - debt_2007)), hjust = -5) +
    scale_x_continuous(labels = scales::dollar_format()) +
    expand_limits(x = 5000) +
    annotate("text", label = "2007", x = 8000, y = 49.8, color = "red", size = 8) +
    annotate("text", label = "2017", x = 19400, y = 49.8, color = "blue", size = 8) +
    labs(x=NULL, 
         y=NULL,
         title="Which States Have Highest Debt Difference?",
         subtitle = "Between Year 2007 and Year 2017") +
   theme(title = element_text(size = 20),
        axis.text.x = element_text(face = "bold", color = "black", size = 15),
        axis.text.y = element_text(color = "black", size = 15),
        legend.text = element_text(color = "black", size = 15),
        panel.grid.minor=element_blank(),
        panel.grid.major.y=element_blank(),
        panel.grid.major.x=element_line(),
        axis.ticks=element_blank(),
        legend.position="top",
        panel.border=element_blank())
```



```{r import-data}
# UGDS: total enrollments

major_data <- read_csv("~/Google Drive/Computer Science MSU/Spring 2019/Data Science/final-project/majors.csv")

merged_data <- read_csv("~/Downloads/CollegeScorecard_Raw_Data/MERGED2016_17_PP.csv") %>%
    select(UNITID, ADM_RATE, AGE_ENTRY, FEMALE, LATITUDE, LONGITUDE, TUITIONFEE_IN, TUITIONFEE_OUT)

scorecard_elements <- read_csv("~/Google Drive/Computer Science MSU/Spring 2019/Data Science/final-project/Most-Recent-Cohorts-Scorecard-Elements.csv") %>% 
    select(UNITID, CITY, STABBR, SAT_AVG, 
           ACTCM25, ACTCM75, ACTCMMID, 
           starts_with("PCIP"), 
           UGDS, GRAD_DEBT_MDN_SUPP)

# names(scorecard_elements)

treasury_elements <- read_csv("~/Google Drive/Computer Science MSU/Spring 2019/Data Science/final-project/Most-Recent-Cohorts-Treasury-Elements.csv") %>%
    select(UNITID, INSTNM, starts_with("PCT_"), 
           MEDIAN_HH_INC, POVERTY_RATE, UNEMP_RATE, 
           PCT25_EARN_WNE_P10, PCT75_EARN_WNE_P10, PCT90_EARN_WNE_P10,
           MN_EARN_WNE_P10, MD_EARN_WNE_P10, 
           MN_EARN_WNE_MALE0_P10, MN_EARN_WNE_MALE1_P10)

college_data <- left_join(scorecard_elements, treasury_elements, by = "UNITID")
college_data <- left_join(college_data, merged_data, by = "UNITID")

# Only consider US states
college_data <- college_data %>%
    filter(STABBR %in% state.abb)

college_data_majors <- college_data %>%
    gather(starts_with("PCIP"), key = "major_id", value = "pct_awarded")

college_data_majors <- left_join(college_data_majors, major_data, by = "major_id")

# college_data %>%
#     count(major_id, major_category)
```

```{r data-clearning}
# convert debt to numbers
# scorecard_elements_cleaned <- scorecard_elements %>%
#     mutate(GRAD_DEBT_MDN_SUPP = as.numeric(GRAD_DEBT_MDN_SUPP))

# add state names and divisions
college_data <- add_column(college_data, 
                           STATENAME = state.name[match(college_data$STABBR, state.abb)], 
                           .after = "STABBR")

college_data <- add_column(college_data, 
                           DIVISION = state.division[match(college_data$STABBR, state.abb)], 
                           .after = "STATENAME")

college_data <- add_column(college_data, 
                           REGION = state.region[match(college_data$STABBR, state.abb)], 
                           .after = "DIVISION")

# convert character columns to numeric
college_data <- college_data %>%
    mutate_at(vars(-UNITID:-REGION, -INSTNM), as.numeric)

college_data <- college_data %>% 
    filter(!STABBR %in% excluded_states)

# add county name
college_data_processed <- left_join(college_data, us_map_data %>% 
                                    select(CITY, STATENAME, county_name), 
                                    by = c("CITY", "STATENAME")) %>%
    group_by(STATENAME) %>%
    mutate_at(vars(-UNITID:-REGION, -INSTNM), 
              ~ifelse(is.na(.x) & is.numeric(.x), median(.x, na.rm = TRUE), .x))

# replace all na values with mean values
# college_data_processed <- college_data_processed %>%
#     group_by(STATENAME) %>%
#     mutate_at(vars(-UNITID:-REGION, -INSTNM), ~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x))

# college_data_majors <- college_data_majors %>% 
#     mutate_at(vars(-UNITID:-REGION, -INSTNM, -major_id, -major_category), as.numeric)

# scorecard_elements_cleaned %>%
#     distinct(STATENAME, DIVISION) %>% View()
```


## What states have least/most debt?
```{r}
college_data_debt <- college_data %>%
    group_by(STATENAME, DIVISION, REGION) %>%
    summarise(median_debt = median(GRAD_DEBT_MDN_SUPP, na.rm = TRUE)) %>%
    arrange(desc(median_debt)) %>%
    ungroup(STATENAME) %>%
    mutate(STATENAME = fct_reorder(STATENAME, median_debt)) 

temp <- brewer.pal(n = 4, name = "Set1")
# temp <- wes_palette(n = 4, name = "GrandBudapest2")
colors = c("Northeast" = temp[1], "South" = temp[2], "North Central" = temp[3], "West" = "#f9c64d")
# colors = c("Northeast" = "#66c2a5", "South" = "#3288bd", "North Central" = "#d73027", "West" = "#d1ff07")

college_data_debt %>%
    head(10) %>%
    ggplot(aes(STATENAME, median_debt, fill = REGION)) +
    geom_col() +
    # geom_text(aes(y = median_debt + 50, label = scales::dollar(median_debt, accuracy = 2)), 
    #           hjust = 0, vjust = 0, 
    #           check_overlap = TRUE) +
    scale_y_continuous(labels = scales::dollar_format()) +
    scale_fill_manual(values = colors) +
    # scale_fill_viridis_d() +
    # coord_flip() +
    coord_polar(start = 0) +
    labs(title = "Top 10 States with Most Debt after 10 years of graduation",
         subtitle = "School Year 2017-2018",
         x = NULL,
         y = "Median Debt",
         fill = NULL) +
    theme(title = element_text(size = 20),
          axis.text.x = element_text(face = "bold", color = "black", size = 15),
          axis.text.y = element_text(face = "bold", color = "black", size = 15),
          legend.text = element_text(color = "black", size = 15))

college_data_debt %>%
    tail(10) %>%
    mutate(STATENAME = fct_reorder(STATENAME, -median_debt)) %>%
    ggplot(aes(STATENAME, median_debt, fill = REGION)) +
    geom_col() +
    scale_y_continuous(labels = scales::dollar_format(), breaks = seq(0, 13000, 2000)) +
    scale_fill_manual(values = colors) +
    coord_flip() +
    # coord_polar() +
    labs(title = "Top 10 States with Least Debt after 10 years of graduation",
         subtitle = "School Year 2017-2018",
         x = NULL,
         y = "Median Debt",
         fill = NULL) +
    theme(title = element_text(size = 20),
          axis.text.x = element_text(face = "bold", color = "black", size = 15),
          axis.text.y = element_text(face = "bold", color = "black", size = 15),
          legend.text = element_text(color = "black", size = 15),
          legend.position = "bottom")

college_data_debt %>%
    ggplot(aes(DIVISION, median_debt, fill = DIVISION)) +
    geom_boxplot(show.legend = FALSE) +
    scale_y_continuous(labels = scales::dollar_format()) +
    # expand_limits(y = 1600) +
    coord_flip() +
    labs(title = "Median Debt among Divisions in the US",
         subtitle = "School Year 2017-2018",
         x = NULL,
         y = "Debt") +
    theme(title = element_text(size = 20),
        axis.text.x = element_text(face = "bold", color = "black", size = 15),
        axis.text.y = element_text(face = "bold", color = "black", size = 15),
        legend.text = element_text(color = "black", size = 15))
```

```{r}
# Create dataset
data=college_data_debt %>%
    transmute(individual = STATENAME,
              group = REGION,
              value = median_debt)

# Set a number of 'empty bar' to add at the end of each group
empty_bar=4
to_add = data.frame( matrix(NA, empty_bar*nlevels(data$group), ncol(data)) )
colnames(to_add) = colnames(data)
to_add$group=rep(levels(data$group), each=empty_bar)
data=rbind(data, to_add)
data=data %>% arrange(group, value)
data$id=seq(1, nrow(data))

# Get the name and the y position of each label
label_data=data
number_of_bar=nrow(label_data)
angle= 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust<-ifelse( angle < -90, 1, 0)
label_data$angle<-ifelse(angle < -90, angle+180, angle)

# Make the plot
ggplot(data, aes(x=as.factor(id), y=value, fill=group)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  geom_bar(stat="identity", alpha=0.8) +
  ylim(-10000,30000) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar() + 
  geom_text(data=label_data, aes(x=id, y=value+1000, label=individual, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=2.5, angle= label_data$angle, inherit.aes = FALSE ) 
```


```{r coord-polar-debt-states}
data <- college_data_debt %>%
    transmute(individual = STATENAME,
              group = REGION,
              value = median_debt)

data = data %>% arrange(group, value)

# Set a number of 'empty bar' to add at the end of each group
empty_bar=5
to_add = data.frame( matrix(NA, empty_bar*nlevels(data$group), ncol(data)) )
colnames(to_add) = colnames(data)
to_add$group=rep(levels(data$group), each=empty_bar)
data=rbind(data, to_add)
data=data %>% arrange(group)
data$id=seq(1, nrow(data))
 
# Get the name and the y position of each label
label_data=data
number_of_bar=nrow(label_data)
angle= 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust<-ifelse( angle < -90, 1, 0)
label_data$angle<-ifelse(angle < -90, angle+180, angle)
 
# prepare a data frame for base lines
base_data=data %>% 
  group_by(group) %>% 
  summarize(start=min(id), end=max(id) - empty_bar) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))
 
# prepare a data frame for grid (scales)
grid_data = base_data
grid_data$end = grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start = grid_data$start - 1
grid_data=grid_data[-1,]

# grid_data_copy <- grid_data %>% add_row(group = "Temp", start = 2, end = 0, title = 12.4) %>% head(3)
 
# Make the plot
ggplot(data, aes(x=as.factor(id), y=value, fill=group)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  
  geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=1) +
  
  # Add a val=100/75/50/25 lines. I do it at the beginning to make sur barplots are OVER it.
  geom_segment(data=grid_data, aes(x = end, y = 25000, xend = start, yend = 25000), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 20000, xend = start, yend = 20000), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 15000, xend = start, yend = 15000), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 10000, xend = start, yend = 10000), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 5000, xend = start, yend = 5000), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  
  # Add text showing the value of each 100/75/50/25 lines
  annotate("text", x = rep(max(data$id),5), y = c(5000, 10000, 15000, 20000, 25000), label = c("$5,000", "$10,000", "$15,000", "$20,000", "$25,000") , color="black", alpha = 0.65, size=4 , angle=0, fontface="bold", hjust=1.1) +
  
  geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=1) +
    scale_fill_manual(values = colors) +
  ylim(-12000,30000) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar() + 
  geom_text(data=label_data, aes(x=id, y=value+1000, label=individual, hjust=hjust), color="black", fontface="bold",alpha=0.8, size=5, angle= label_data$angle, inherit.aes = FALSE ) +
  
  # Add base line information
  geom_segment(data=base_data, aes(x = start, y = -5, xend = end, yend = -5), colour = "black", alpha=0.8, size=0.6 , inherit.aes = FALSE )  +
  geom_text(data=base_data, aes(x = title, y = -18, label=group), hjust=c(1,1,0,0), colour = "black", alpha=0.8, size=4, fontface="bold", inherit.aes = FALSE)

data %>%
    filter(!is.na(individual)) %>%
    ggplot(aes(individual, value, fill = group)) +
    geom_col() +
    scale_fill_manual(values = colors) +
    scale_y_continuous(labels = scales::dollar_format()) +
    coord_flip() +
    facet_wrap(~ group, scales = "free_y") +
    labs(y = NULL,
         x = NULL) +
    theme(axis.text.x = element_text(color = "black", size = 15),
        axis.text.y = element_text(face = "bold", color = "black", size = 20),
        legend.text = element_text(color = "black", size = 20),
           strip.text = element_text(size = 15))
```

## Standardized Test
Let's look at standardized test scores both in __SAT__ and __ACT__.

### Average SAT Score within divisions
```{r}
p <- college_data %>%
    ggplot(aes(DIVISION, SAT_AVG, fill = DIVISION)) +
    geom_boxplot(show.legend = FALSE) +
    scale_y_continuous(labels = scales::comma_format(), breaks = seq(0, 1800, 200)) +
    expand_limits(y = 1600) +
    coord_flip() +
    labs(title = "Average SAT Scores within Divisions in the US",
         subtitle = "School Year 2017-2018",
         x = NULL,
         y = "Score") +
    theme(title = element_text(size = 20),
          axis.text.x = element_text(face = "bold", color = "black", size = 20),
          axis.text.y = element_text(face = "bold", color = "black", size = 20),
          legend.text = element_text(color = "black", size = 20)) + 
    transition_states(DIVISION,
                      transition_length = 2,
                      state_length = 1) +
    enter_fade() +
    exit_shrink() +
    shadow_mark() +
    ease_aes('sine-in-out')


animate(p, duration = 7, height = 1000, width = 1600, renderer = gifski_renderer(loop = FALSE))

anim_save(filename = "box_plot.gif", path = "~/Documents/")
```

### 25th and 75th percentile ACT cumulative score
```{r}
college_data %>%
    group_by(REGION, DIVISION) %>%
    summarise(ACTCMMID = median(ACTCMMID, na.rm = TRUE),
              ACTCM25 = median(ACTCM25, na.rm = TRUE),
              ACTCM75 = median(ACTCM75, na.rm = TRUE)) %>%
    mutate(DIVISION = fct_reorder(DIVISION, ACTCM75)) %>%
    ggplot(aes(DIVISION, ACTCMMID, color = as.factor(REGION))) +
    geom_point(size = 5) +
    geom_errorbar(aes(ymin = ACTCM25, ymax = ACTCM75), size = 2, width = 0.5) +
    scale_color_brewer(palette = "Dark2") +
    coord_flip() +
    scale_y_continuous(breaks = seq(18, 30, 2)) +
    expand_limits(y = 18) +
    labs(title = "ACT score distribution over divisions and regions",
         subtitle = "25th percentile, median, and 75th percentile scores",
         x = NULL,
         y = "Score",
         color = NULL) +
    theme(title = element_text(size = 20),
          axis.text.x = element_text(face = "bold", color = "black", size = 15),
          axis.text.y = element_text(face = "bold", color = "black", size = 15),
          legend.text = element_text(color = "black", size = 15),
          legend.position = "bottom")
```

### Some histograms

Histogram on debt and on enrollment

```{r}
college_data %>%
    ggplot(aes(GRAD_DEBT_MDN_SUPP)) +
    geom_histogram(bins = 50, fill = "#4286f4") +
    scale_x_continuous(labels = scales::dollar_format())

college_data %>%
    filter(UGDS > 1000) %>%
    ggplot(aes(UGDS)) +
    geom_density(fill = "#4286f4") +
    scale_x_log10()

# college_data %>%
#     count(UGDS, sort = TRUE) %>% View()
```

## Correlations
This section examines correlations between some continuous variables in the dataset
### Correlation between ACT and SAT
```{r}
college_data %>%
    ggplot(aes(ACTCMMID, SAT_AVG)) +
    geom_hex() +
    coord_cartesian() +
    facet_wrap(REGION ~ .) +
    labs(title = "Will Students Score High in ACT also Score High in SAT?",
         subtitle = "Is there any difference of performance between regions?",
         x = "ACT Score",
         y = "SAT Score") +
    theme(title = element_text(color = "black", size = 20),
          legend.text = element_text(color = "black", size = 20),
          axis.text.x = element_text(face = "bold", color = "black", size = 20),
          axis.text.y = element_text(face = "bold", color = "black", size = 20),
          strip.text = element_text(size = 20))
```

### Correlation between __Admission Rate__ and __Median Earning__ After 10 Years
```{r}
top_earnings <- college_data %>%
    filter(ADM_RATE > 0) %>%
    arrange(desc(PCT90_EARN_WNE_P10)) %>%
    head(10)

library(ggExtra)

g <- college_data %>%
    filter(ADM_RATE > 0) %>%
    ggplot(aes(ADM_RATE, PCT90_EARN_WNE_P10)) +
    geom_point(aes(color = REGION), size = 3) +
    geom_encircle(data=top_earnings, aes(ADM_RATE, PCT90_EARN_WNE_P10), color = "red", size = 2, expand = 0.08) +
    geom_smooth(aes(color = REGION), method = "lm", size = 2, se = FALSE) +
    scale_x_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
    scale_y_continuous(labels = scales::dollar_format()) +
    facet_wrap(~ REGION, scales = "free_x") +
    labs(title = "Will Admission Rate Determine The Salary After Graduation?",
         x = "Admission Rate", 
         y = NULL,
         color = NULL) +
    theme(title = element_text(color = "black", size = 20),
          legend.text = element_text(color = "black", size = 20),
          strip.text = element_text(size = 20),
          axis.text.x = element_text(face = "bold", color = "black", size = 15),
          axis.text.y = element_text(face = "bold", color = "black", size = 15),
          legend.position = "bottom")

# p + transition_states(
#         REGION, transition_length = 2,
#         state_length = 1
#     ) +
#     enter_fade() +
#     exit_shrink() +
#     ease_aes('sine-in-out')
# 
# p
# animate(p)

g

ggMarginal(g, type = "histogram", fill = "transparent")

college_data %>%
    ggplot(aes(ACTCMMID)) +
    geom_density(aes(fill = factor(REGION)), alpha = 0.5) +
    labs(title="Student ACT Scores Density plot", 
         subtitle="Grouped By Regions",
         x="Score",
         y = NULL,
         fill="Region") +
    theme(title = element_text(color = "black", size = 20),
          legend.text = element_text(color = "black", size = 20),
          strip.text = element_text(size = 20),
          axis.text.x = element_text(face = "bold", color = "black", size = 15),
          axis.text.y = element_text(face = "bold", color = "black", size = 15),
          legend.position = "bottom")
```

### Correlation between __Admission Rate__ and __Unemployment Rate__ After 10 Years
```{r}
college_data %>%
    ggplot(aes(UNEMP_RATE / 100, color = REGION)) +
    geom_freqpoly(binwidth = 0.01) +
    scale_x_continuous(labels = scales::percent_format())
```


```{r}
college_data %>%
    ggplot(aes(TUITIONFEE_IN, MD_EARN_WNE_P10)) +
    geom_point(aes(color = REGION), alpha = 0.75, show.legend = FALSE) +
    scale_x_continuous(labels = scales::dollar_format()) +
    scale_y_continuous(labels = scales::dollar_format()) +
    facet_wrap(REGION ~ .) +
    labs(title = "Will Costly Schools Help Students Get Higher-Earning Jobs?",
         subtitle = "Median salaries are collected 10 years after graduation",
         caption = "School Year 2017-2018",
         x = "In-state Tuition Fee",
         y = NULL) +
    theme(title = element_text(color = "black", size = 20),
          legend.text = element_text(color = "black", size = 20),
          axis.text.x = element_text(face = "bold", color = "black", size = 15),
          axis.text.y = element_text(face = "bold", color = "black", size = 15),
          strip.text = element_text(size = 20))

# college_data %>%
#     filter(!is.na(MD_EARN_WNE_P10)) %>%
#     arrange(desc(MD_EARN_WNE_P10)) %>% 
#     select(TUITIONFEE_IN, TUITIONFEE_OUT, ADM_RATE, MD_EARN_WNE_P10) %>%
#     View()

# college_data %>%
#     filter(is.na(TUITIONFEE_IN)) %>%
#     count()
```

## What majors earn the most?
```{r}
college_data %>%
    gather(MN_EARN_WNE_MALE0_P10, MN_EARN_WNE_MALE1_P10, key = "gender", value = "gender_earning") %>% group_by(STATENAME) %>%
    summarise(median_salary = median())
```


### playing with maps

Earnings from all colleges in the US
```{r}

income_college_data <- college_data %>%
    filter(!is.na(MD_EARN_WNE_P10) & !is.na(INSTNM)) %>%
    distinct(INSTNM, .keep_all = TRUE)

 income_college_data %>%
    ggplot(aes(LONGITUDE, LATITUDE, size = MD_EARN_WNE_P10, color = MD_EARN_WNE_P10)) + 
    borders("state") +
    geom_point(alpha = 0.8) +
    scale_color_continuous(labels = scales::dollar_format(),
                           low = "#132B43", high = "#56B1F7") +
    scale_size(labels = scales::dollar_format()) +
    coord_map("albers", parameters = c(39, 45)) + 
    labs(title = "Median Salary after 10 Years of Graduation",
          x = NULL,
          y = NULL,
          color = NULL,
          size = "Median Salary") +
    theme(title = element_text(size = 20),
          legend.text = element_text(color = "black", size = 20))
```

### Earnings of some Southern States
```{r income-college-south}
income_college_data_south <- income_college_data %>% 
    filter(STABBR %in% c("MS", "AL", "TN", "LA", "AR"))

state_map_data <- map_data("state")

income_college_data_south %>%
    arrange(MD_EARN_WNE_P10) %>%
    ggplot(aes(LONGITUDE, LATITUDE, 
                size = MD_EARN_WNE_P10, 
                color = MD_EARN_WNE_P10)) +
    borders("state", c("mississippi", "alabama", "tennesse", "louisiana", "arkansas"), colour = "black") +
    # geom_map(data = state_map_data, map = subset(state_map_data, region %in% c("mississippi", "alabama", "tennesse", "louisiana")), aes(map_id = group), fill = NA, color = "pink") +
    geom_point(alpha = 1, position = "jitter") +
    # geom_label(data = income_college_data_south %>% arrange(MD_EARN_WNE_P10) %>% tail(3), aes(label = INSTNM), nudge_x = 2, nudge_y = 2) +
    geom_label_repel(data=income_college_data_south %>% arrange(MD_EARN_WNE_P10) %>% tail(3),
                    aes(label = INSTNM), size = 8,
                    nudge_x = 2, nudge_y = 4.5) +
    geom_label_repel(data=income_college_data_south %>% arrange(MD_EARN_WNE_P10) %>% head(3),
                    aes(label = INSTNM), size = 8,
                    nudge_x = 4, nudge_y = -6,
                    max.iter = 20000) +
    scale_color_continuous(labels = scales::dollar_format(),
                           breaks = seq(25000, 125000, 25000),
                           low = "red", high = "green") +
    # scale_colour_gradient2(low = "red",
    #                        mid = scales::muted("red"),
    #                        high = "green",
    #                        space = "Lab",
    #                        labels = scales::dollar_format(),
    #                        breaks = seq(25000, 125000, 25000)) +
    scale_size(range = c(0, 10),
               breaks = seq(25000, 125000, 25000), guide = "none") +
    coord_map() + 
    ylim(28, 40) +
    # xlim(-95, -80) +
    labs(title = "What Southern States That Graduates Make Most Money?",
         subtitle = "Tennesse, Arkansas, Mississippi, Alabama, Louisiana",
         x = NULL,
         y = NULL,
         color = "Median Salary",
         size = NULL) +
    theme(title = element_text(size = 20),
          legend.text = element_text(color = "black", size = 20, margin = margin(t = 10)),
          legend.key.height = unit(2, "lines")) +
    # theme_void() +
    annotate("text", x = -89.5, y = 33, label = "MS", size = 12) +
    annotate("text", x = -92.5, y = 31, label = "LA", size = 12) +
    annotate("text", x = -86.5, y = 33, label = "AL", size = 12) +
    annotate("text", x = -86, y = 35.5, label = "TN", size = 12) +
    annotate("text", x = -92.5, y = 35.7, label = "AR", size = 12)
```

## Chropleth maps
Plot 90th percentile median salary of each state
```{r}
us_states <- map_data("state")

us_states_abb <- tibble(
    abb = state.abb,
    long = state.center$x,
    lat = state.center$y
)


# ggplot(us_states, aes(long, lat, group = group, fill = region)) +
#     geom_polygon(color = "grey90", size = 0.1) +
#     coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
#     guides(fill = FALSE)

median_salary_states <- college_data %>%
    filter(!is.na(PCT90_EARN_WNE_P10)) %>%
    group_by(STATENAME, STABBR) %>%
    summarise(median_salary = median(PCT90_EARN_WNE_P10)) %>%
    # summarise(median_salary = sum(PCT90_EARN_WNE_P10, na.rm = TRUE) / n()) %>%
    mutate(region = str_to_lower(STATENAME))

salary_us_state <- left_join(median_salary_states, us_states, by = "region")

ggplot(data = salary_us_state, aes(long, lat, group = group, fill = median_salary)) +
    geom_polygon(color = "grey90", size = 0.1) +
    geom_text(data = us_states_abb, aes(long, lat, label = abb), inherit.aes = FALSE) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    # scale_fill_brewer(palette = "Blues") +
    # scale_fill_continuous(labels = scales::dollar_format()) +
    scale_fill_viridis_c(labels = scales::dollar_format()) +
    # guides(fill = guide_legend(nrow = 1)) +
    xlim(-125, -70) +
    theme_map() +
    labs(title = "Which states have highest-earning graduates?",
         subtitle = "Salaries calculated by median of 90th percentile all salaries",
         caption = "Data collected in 2018",
         fill = NULL) +
    theme(title = element_text(size = 20),
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 20),
          legend.position = "bottom",
          legend.justification = "center",
          legend.key.width = unit(1.25, "in"))
# 
# ggplotly(g)

#----- Unemployment Rates accross States ---- #
unemployment_rate_states <- college_data_processed %>%
    group_by(STATENAME, STABBR) %>%
    summarise(mean_unemployment_rate = mean(UNEMP_RATE)) %>%
    # summarise(median_salary = sum(PCT90_EARN_WNE_P10, na.rm = TRUE) / n()) %>%
    mutate(region = str_to_lower(STATENAME))

unemployment_us_state <- left_join(unemployment_rate_states, us_states, by = "region")

ggplot(data = unemployment_us_state, aes(long, lat, group = group, fill = mean_unemployment_rate / 100)) +
    geom_polygon(color = "grey90", size = 0.1) +
    geom_text(data = us_states_abb, aes(long, lat, label = abb), inherit.aes = FALSE) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    # scale_fill_brewer(palette = "Blues") +
    # scale_fill_continuous(labels = scales::dollar_format()) +
    scale_fill_viridis_c(labels = scales::percent_format()) +
    # guides(fill = guide_legend(nrow = 1)) +
    xlim(-125, -70) +
    theme_map() +
    labs(title = "Which states have highest unemployment rate?",
         caption = "Data collected in 2018",
         fill = NULL) +
    theme(title = element_text(size = 20),
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 20),
          legend.position = "bottom",
          legend.justification = "center",
          legend.key.width = unit(1.25, "in"))

college_data_majors %>%
    filter(STABBR == "MT") %>%
    group_by(major_category) %>%
    mutate(pct_awarded = as.numeric(pct_awarded)) %>%
    summarise(pct_graduate = mean(pct_awarded, na.rm = TRUE)) %>%
    mutate(major_category = fct_reorder(major_category, pct_graduate)) %>%
    ggplot(aes(major_category, pct_graduate)) +
    geom_col() +
    scale_y_continuous(labels = scales::percent_format()) +
    coord_flip() +
    labs(x = NULL,
         y = NULL)
```

## Animation for unemployment rates over 20 years (1998-2017)
```{r}
# year_state_unemployment <- processed_data_by_years %>% 
#     group_by(YEAR, state_id, STATENAME) %>%
#     summarise(mean_unemployment_rate = mean(UNEMP_RATE, na.rm = TRUE)) %>%
#     mutate(region = str_to_lower(STATENAME)) %>%
#     ungroup() %>% filter(YEAR < 2007)
# 
# year_state_unemployment_map <- left_join(year_state_unemployment, us_states, by = "region")
# 
# g <- ggplot(data = year_state_unemployment) +
#     geom_polygon(data=year_state_unemployment_map, aes(long, lat, group = group, 
#                      fill = mean_unemployment_rate / 100), 
#                  color = "grey90", size = 0.1) +
#     geom_text(data = us_states_abb, aes(long, lat, label = abb), inherit.aes = FALSE) +
#     coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
#     # scale_fill_brewer(palette = "Blues") +
#     # scale_fill_continuous(labels = scales::dollar_format()) +
#     scale_fill_viridis_c(labels = scales::percent_format()) +
#     # guides(fill = guide_legend(nrow = 1)) +
#     xlim(-125, -70) +
#     theme_map() +
#     labs(title = "Year = {frame_time}",
#          caption = "Data collected in 2018",
#          fill = NULL) +
#     theme(title = element_text(size = 20),
#           legend.text = element_text(size = 20),
#           legend.title = element_text(size = 20),
#           legend.position = "bottom",
#           legend.justification = "center",
#           legend.key.width = unit(1.25, "in")) 
# 
# g
# 
# # g1 <- g + transition_time(YEAR) +
# #     ease_aes('linear')
# # 
# # animate(g1, interval = .2, nframes = 20, fps = 10)
# 
# # ggplotly(g)
# 
# ggplot(year_state_unemployment, aes(mean_unemployment_rate)) +
#     geom_histogram(bins = 30) +
#     transition_time(YEAR) +
#     ease_aes('linear') +
#     labs(title = "Year = {frame_time}")
```


## 90 Percentile salary accross the counties
```{r}
us_counties <- map_data("county")

median_salary_counties <- college_data_processed %>%
    group_by(county_name) %>%
    summarise(median_salary = median(PCT90_EARN_WNE_P10)) %>%
    mutate(subregion = str_to_lower(county_name))

salary_us_county <- left_join(us_counties, median_salary_counties, by = "subregion")

ggplot(data = salary_us_county, aes(long, lat, group = group, fill = median_salary)) +
    geom_polygon(color = "grey90", size = 0.1) +
    # geom_text(data = us_states_abb, aes(long, lat, label = abb), inherit.aes = FALSE) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    # scale_fill_brewer(palette = "Blues") +
    # scale_fill_continuous(labels = scales::dollar_format()) +
    scale_fill_viridis_c(labels = scales::dollar_format()) +
    # guides(fill = guide_legend(nrow = 1)) +
    xlim(-125, -70) +
    theme_map() +
    labs(title = "Which counties have highest-earning graduates?",
         subtitle = "Salaries calculated by median of 90th percentile all salaries",
         caption = "Data collected in 2018",
         fill = NULL) +
    theme(title = element_text(size = 20),
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 20),
          legend.position = "bottom",
          legend.justification = "center",
          legend.key.width = unit(1.25, "in"))

median_salary_counties %>% arrange(desc(median_salary))

college_data_processed %>%
    filter(STABBR == "MS") %>% View()

college_data_processed %>%
    arrange(desc(PCT90_EARN_WNE_P10)) %>% select(
        CITY, STATENAME, PCT90_EARN_WNE_P10, INSTNM, county_name
    ) %>% View()

college_data_processed %>%
    filter(county_name == "San Francisco") %>% select(
        CITY, STATENAME, PCT90_EARN_WNE_P10, INSTNM, county_name
    ) %>% View()
```



## Modeling
Predict salary based on __admission rate__, __SAT score__, __ACT score__

```{r}
college_data_by_state <- college_data_processed %>%
    group_by(STATENAME, DIVISION, REGION) %>%
    nest()

salary_state_model <- function(df) {
    lm(MD_EARN_WNE_P10 ~ ADM_RATE + SAT_AVG + ACTCMMID, data = df)
}

college_data_by_state <- college_data_by_state %>%
    mutate(model = purrr::map(college_data_by_state$data, salary_state_model))

college_data_by_state <- college_data_by_state %>%
    mutate(resids = purrr::map2(data, model, add_residuals))

resids <- unnest(college_data_by_state, resids)

glance <- college_data_by_state %>%
    mutate(glance = purrr::map(model, broom::glance)) %>%
    unnest(glance, .drop = TRUE)

# resids %>%
#     ggplot(aes())
```

```{r}
model <- lm(MD_EARN_WNE_P10 ~ SAT_AVG, data=college_data)

grid <- college_data_processed %>%
    data_grid(SAT_AVG = seq_range(SAT_AVG, 30)) %>%
    add_predictions(model, "MD_EARN_WNE_P10")

ggplot(data=college_data_processed, aes(SAT_AVG, MD_EARN_WNE_P10)) +
    geom_hex(bins = 50) +
    geom_line(data=grid, color="red", size=1)

college_data_processed %>%
    filter(STABBR == "WI") %>%
    View()

college_data %>%
    add_residuals(model, "lresid") %>%
    ggplot(aes(SAT_AVG, lresid)) +
    geom_hex(bins = 50) +
    geom_ref_line(h = 0, colour = "red")
```



